Data Analysis Extensions

### 2015-10-15 RFCs/HDF5_Library/AnalysisExt/analysis_ext-body.tex
A few fixes to previous commit
Former-commit-id: 66a36b3bc72dbb39f4f492b15a12021995ef594e
Update analysis RFC with cross-container queries and add latexdiff support
Former-commit-id: a8e6bf95a7c3ec932a96439f46463b0f574ef90b

\section{Index Objects}
Index objects are designed to accelerate creation of views from query operations.
For example, if the ``$\mathrm{(link\_name = Pressure)}$
$\mathrm{\land}$ $\mathrm{(data\_element = 17)}$'' query
(previously described) is going to be either frequently executed or executed
on a container that contains significantly large datasets, indices
could be created in that container, which would speed up the
creation of views when querying for link names and for data element values.
Indices created for accelerating the $link\_name = Pressure$ or $data\_element = 17$ queries
would also improve view creation for the more complex
$(link\_name = Pressure)\land(data\_element = 17)$ query.
Although creating indices for metadata components of queries, such as link or
attribute names, is possible, we first focus on index creation for dataset elements,
as they represent the largest volume of data in typical HPC application usage of
HDF5. Queries with metadata components execute properly,
but are not able to be accelerated with an index currently.

@@

{\lstsetc
\begin{lstlisting}
herr_t H5Dquery(hid_t dset_id, hid_t space_id, hid_t query_id, hid_t xapl_id,
    hid_t *space_id);
\end{lstlisting}
}

@@

\subsection{Support for HDF5 Compound Types}

@@

\subsection{Support for HDF5 Chunking}

@@

\subsection{Support for Parallel Indexing}

@@

    /* Use query to select elements in the dataset. */
    result_space_id = H5Dquery(dataset_id, query_id);

@@

    /* Use result. */

>>>>

\subsection{Cross Container Queries}

In the case where multiple containers (files, locations) need to be queried,
a single query operation may be used instead of performing individual queries
on each container. Performing queries across multiple containers with a single
operation can enable analysis operations to generate an aggregated
view of query results more easily, as well as enable a silent execution of the
query in parallel, thereby reducing the time to generate a view from multiple containers.

The first extension to the previously described \texttt{H5Qapply} call is defined
with the following routine, \texttt{H5Qapply\_multi}:

{\lstsetc
\begin{lstlisting}
hid_t H5Qapply_multi(size_t loc_count, hid_t *loc_ids, hid_t query_id,
  unsigned *result, hid_t vcpl_id);
\end{lstlisting}
}

Location IDs specified may be entire containers, groups in a container's hierarchy
or individual datasets within a container. These locations may reside within the
same container, or in multiple containers. Additionally, the \texttt{H5Qapply\_str}
routine allows the user to provide container names instead of HDF5 locations:

{\lstsetc
\begin{lstlisting}
hid_t H5Qapply_str(size_t cont_count, const char *cont_name[], hid_t query_id,
  unsigned *result, hid_t vcpl_id);
\end{lstlisting}
}

As a direct consequence, one may want to pass queries that match a particular
container name. Therefore the \texttt{H5Q\_TYPE\_CONT\_NAME} query type can be
used to construct a query that matches this new condition. The matching
condition may be passed in the form of a file name or a regular
expression.

(Optional?) Apply query on all the containers of a given directory with \texttt{H5Qapply\_dir}
{\lstsetc
\begin{lstlisting}
hid_t H5Qapply_dir(const char *path, hid_t query_id, unsigned *result, hid_t vcpl_id);
\end{lstlisting}
}

Applying cross-container queries creates a view that contains references
to external locations (as opposed to the previous \texttt{H5Qapply} call where
references were internal). Therefore, the extended HDF5 reference object
also includes a container identifier so that, when the view result is
given back to the user, a subsequent call to \texttt{H5Rdereference} can return
a reference to an object within an externally referenced container\footnote{
As a result, the HDF5 reference API does not require any changes.}.

Parallel execution of the query optional? \textit{Pass hint for parallel execution
through property list.}

\section{Index Objects}
Index objects are designed to accelerate creation of views from query operations.
For example, if the ``$\mathrm{(link\_name = Pressure)}$
$\mathrm{\land}$ $\mathrm{(data\_element = 17)}$'' query
(previously described) is going to be either frequently executed or executed
on a container that contains significantly large datasets, indices
could be created in that container, which would speed up the
creation of views when querying for link names and for data element values.
Indices created for accelerating the  ``$\mathrm{(link\_name = Pressure)}$'' or
``$\mathrm{(data\_element}$ $\mathrm{= 17)}$'' queries
would also improve view creation for the more complex
``$\mathrm{(link\_name = Pressure)}$ $\mathrm{\land}$
$\mathrm{(data\_element = 17)}$'' query.
Although creating indices for metadata components of queries, such as link or
attribute names, is possible, we first focus on index creation for dataset elements,
as they represent the largest volume of data in typical HPC application usage of
HDF5. Queries with metadata components execute properly,
but are not able to be accelerated with an index currently.

@@

{\lstsetc
\begin{lstlisting}
hid_t H5Dquery(hid_t dset_id, hid_t space_id, hid_t query_id, hid_t xapl_id);
\end{lstlisting}
}

@@

\subsection{Support for HDF5 Compound Types (TBD)}

@@

\subsection{Support for HDF5 Chunking (TBD)}

@@

\subsection{Support for Parallel Indexing (TBD)}

@@

    /* Use query to select elements in the dataset. */
    result_space_id = H5Dquery(dataset_id, H5S_ALL, query_id, H5P_DEFAULT);

@@

    /* Use result. */
    [...]

### 2015-10-15 RFCs/HDF5_Library/AnalysisExt/analysis_ext.tex
Update analysis RFC with cross-container queries and add latexdiff support
Former-commit-id: a8e6bf95a7c3ec932a96439f46463b0f574ef90b

% Code snippets
\usepackage{listings}
\def\lstsetc{\lstset{language=C,
  numbers=left,
  xleftmargin=1.8em,
  frame=L,
  framexleftmargin=0.2em,
  rulecolor=\color{rulecolor},
  numberstyle=\tiny\color{gray},
  stepnumber=1,
  showspaces=false, 
  showstringspaces=false,
  breaklines=true,
  basicstyle=\footnotesize\ttfamily,
  stringstyle=\itshape,
  commentstyle=\itshape\bfseries,
  morekeywords={main, size_t, malloc, free,
    hsize_t, hid_t, herr_t, H5X_class_t, H5X_type_t, H5Q_combine_op_t,
    H5Q_type_t, H5Q_match_op_t,
    H5Fcreate, H5Fclose,
    H5Dcreate, H5Dopen, H5Dclose, H5Dwrite, H5Dread, H5Dquery,
    H5Qcreate, H5Qclose, H5Qcombine, H5Qget_type, H5Qget_match_op,
    H5Qget_components, H5Qget_combine_op, H5Qencode, H5Qdecode,
    H5Qapply,
    H5Screate_simple, H5Sclose, H5Sget_select_npoints,
    H5Xregister, H5Xunregister, H5Xcreate, H5Xremove, H5Xget_count, H5Xget_size,
    H5Xget_info
    }
  }
}

% Title, author, etc
\title{Data Analysis Extensions}
\author{Jerome Soumagne}
\author{Quincey Koziol}
\date{Oct 14, 2015}
\rfcversion{2014-07-17.v6}
\revision{Jul. 17, 2014}{Version 1 circulated for comment within The HDF Group.}
\revision{Aug. 20, 2014}{Version 2 circulated for comment within The HDF Group.}
\revision{Nov. 14, 2014}{Version 3 circulated for comment within The HDF Group.}
\revision{Feb. 25, 2015}{Version 4 edits and typo fixing.}
\revision{Jul. 24, 2015}{Version 5 edits to view object.}
\revision{Oct. 14, 2015}{Version 6 rework of query/view sections.}

>>>>

% Code snippets
\usepackage{listings}
\def\lstsetc{\lstset{language=C,
  %numbers=left,
  xleftmargin=1.8em,
  frame=L,
  framexleftmargin=0.2em,
  rulecolor=\color{rulecolor},
  numberstyle=\tiny\color{gray},
  stepnumber=1,
  showspaces=false, 
  showstringspaces=false,
  breaklines=true,
  basicstyle=\footnotesize\ttfamily,
  stringstyle=\itshape,
  commentstyle=\itshape\bfseries,
  morekeywords={main, size_t, malloc, free,
    hsize_t, hid_t, herr_t,
    H5X_class_t, H5X_type_t, H5X_info_t,
    H5Q_combine_op_t, H5Q_type_t, H5Q_match_op_t,
    H5Fcreate, H5Fclose,
    H5Dcreate, H5Dopen, H5Dclose, H5Dwrite, H5Dread, H5Dquery,
    H5Qcreate, H5Qclose, H5Qcombine, H5Qget_type, H5Qget_match_op,
    H5Qget_components, H5Qget_combine_op, H5Qencode, H5Qdecode,
    H5Qapply, H5Qapply_multi, H5Qapply_str, H5Qapply_dir,
    H5Screate_simple, H5Sclose, H5Sget_select_npoints,
    H5Xregister, H5Xunregister, H5Xcreate, H5Xremove, H5Xget_count, H5Xget_size,
    H5Xget_info
    }
  }
}

% Title, author, etc
\title{Data Analysis Extensions}
\author{Jerome Soumagne}
\author{Quincey Koziol}
\date{Oct 15, 2015}
\rfcversion{2014-07-17.v7}
\revision{Jul. 17, 2014}{Version 1 circulated for comment within The HDF Group.}
\revision{Aug. 20, 2014}{Version 2 circulated for comment within The HDF Group.}
\revision{Nov. 14, 2014}{Version 3 circulated for comment within The HDF Group.}
\revision{Feb. 25, 2015}{Version 4 edits and typo fixing.}
\revision{Jul. 24, 2015}{Version 5 edits to view object.}
\revision{Oct. 14, 2015}{Version 6 rework of query/view sections.}
\revision{Oct. 15, 2015}{Version 7 first draft of cross-container queries.}
